---
title: "Question 12"
author: "samyak"
output: github_document
---

## Program to compute the measures of dispersion for the following frequency distribution.

midpoint | number
--- | ---
95.5 | 5
105.5 | 8
115.5 | 22
125.5 | 27
135.5 | 17
145.5 | 9
155.5 | 5
165.5 | 5
175.5 | 2

### Measures of dispersion?
Measures of dispersion - such as range, variance, standard deviation, quantiles - and can be calculated with standard functions present in the native *stats* package. In addition summary function can be used for other statistical information.

### Storing Data in a Data Frame
```{r}
blood_pressures <- seq(from = 95.5, to = 175.5, by = 10) 
frequency_dist <- c(5,8,22,27,17,9,5,5,2)
bpDist <- data.frame(bp = blood_pressures, freq = frequency_dist)
bpDist
```

### Calculating Range
```{r}
range(bpDist$bp)
```
This shows that the minimum bp value in the distribution is 95.5 and the maximum value is 175.5

### Calculating Variance and Standard Deviation

Since we have a frequency distribution, we have to use weighted mean, variance, and sd to reflect the frequency of each blood pressure measure in the calculation of each of these.

```{r}
bpMean <- weighted.mean(bpDist$bp, bpDist$freq)
bpVar <- sum(bpDist$freq * (bpDist$bp - bpMean)^2) / sum(bpDist$freq)
bpSd <- sqrt(bpVar)
```

**Weighted Mean** of the frequency distribution is
```{r}
bpMean
```

**Weighted Variance** of the frequency distribution is 
```{r}
bpVar
```

**Weighted Standard Deviation** of the frequency distribution is 
```{r}
bpSd
```

### Weighted Quantile

First we get the vector of all values in the frequency distribution.

```{r}
bpVec <- rep(bpDist$bp, times = bpDist$freq)
bpVec
```

Now we can get the different quantile values

```{r}
quantile(bpVec, c(0,0.25,0.5,0.75,1))
```