---
title: "Question 13"
author: "samyak"
output: github_document
---

## Program to calculate probability distribution of a random variable X that is the number of the face on top when a dice is rolled.

## Code

The method that we are using to obtain a probability distribution is the repeated random sampling procedure.

### Getting random sample

Here we use the function `sample` that takes a sample of specified size from elements of a vector with or without replacement.  
So to get a random sample of size 20 of the heads of dice we can use the following

```{r}
dice_numbers <- 1:6
dice_sample <- sample(dice_numbers, size = 20, replace = TRUE)
```

Doing this we get the random sample as
```{r, echo=FALSE}
dice_sample
```

### Approximating probabilities of a random sample

We can count the frequencies of each random outcome using the `table` function

```{r}
table(dice_sample)
```

Now what we need is the *relative* frequency. To obtain that we need to divide the frequency of each outcome by the number of times the experiment was carried out.

```{r}
probability_dist <- table(dice_sample) / length(dice_sample)
probability_dist
```

### Visual representation of the distribution

We can use the `plot` function with the table as input and plot the probability distribution of occurence of each outcome.

```{r}
plot(probability_dist, xlab = "Dice number", ylab = "Probability", main = "sample size = 20")
```


## Examples

### random sample of size 40

```{r}
dice_sample <- sample(dice_numbers, size = 40, replace = TRUE)
probability_dist <- table(dice_sample) / length(dice_sample)
plot(probability_dist, xlab = "Dice number", ylab = "Probability", main = "sample size = 40")
```

### random sample of size 70

```{r}
dice_sample <- sample(dice_numbers, size = 70, replace = TRUE)
probability_dist <- table(dice_sample) / length(dice_sample)
plot(probability_dist, xlab = "Dice number", ylab = "Probability", main = "sample size = 70")
```

### random sample of size 100

```{r}
dice_sample <- sample(dice_numbers, size = 100, replace = TRUE)
probability_dist <- table(dice_sample) / length(dice_sample)
plot(probability_dist, xlab = "Dice number", ylab = "Probability", main = "sample size = 100")
```

### random sample of size 1000

```{r}
dice_sample <- sample(dice_numbers, size = 1000, replace = TRUE)
probability_dist <- table(dice_sample) / length(dice_sample)
plot(probability_dist, xlab = "Dice number", ylab = "Probability", main = "sample size = 1000")
```