---
title: "Question 11"
author: "kaymas"
output: github_document
---

## **Program** To find basic statistical information about students enrolled in a course.

The initial data set given in the question is  
`20 20 20 20 20 21 21 21 22 22 22 22 23 23 23`

```{r echo=FALSE}
age <- c(20,20,20,20,20,21,21,21,22,22,22,22,23,23,23)
```

Now, R provides built in functions for finding mean and median but none for finding mode or modal value. So here we define it ourselves.

```{r definition}
modal <- function(data.vector){
  unique.elements <- unique(data.vector)
  modal.value <- unique.elements[which.max(table(data.vector))]
  return (modal.value)
}
```

Here  
+ `unique` function outputs the unique elements of the data vector.  
+ `which.max` function determines the *location* of the **first** maximum element of the data vector  
+ `table` function in this context calculates the frequency of the levels of the data vector  

### Part - 1
####Median age of all students under 22 years

```{r}
median(age[age < 22])
```

### Part - 2
####Median age of all students

```{r}
median(age)
```

### Part - 3
####Mean age of all students

```{r}
mean(age)
```

### Part - 4
####Modal age for all students

```{r}
modal(age)
```

### Part - 5
####Mean, Median, Modal age of all students when 2 new students, both of age 23, are added to data set.

```{r}
age <- c(age, c(23,23))
mean(age)
median(age)
modal(age)
```


